### 2016-08-08 presentation/index.Rmd
fixes

*

@@

## Variables

Variables According to the HUD Document Data Set

@@

```{r}

@@

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

@@

```{r}

  cleanyearsmonthly %>%  
  group_by_("year","region", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

@@

map_df %>% 
  ggplot() +  
  geom_polygon(aes(x=long, 
                   y = lat, 
                   fill = zinc2_median, 
                   group = group)) + 
  coord_fixed(1.3) +
  theme_tufte() +
  scale_x_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  scale_y_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  labs(fill = "Median income") +
  scale_fill_distiller(limits = c(25000, 42000), 
                                 breaks = sort(unique(map_df$zinc2_median),
                                               decreasing = T),
                                 palette = "Greens", 
                                 direction = 1,
                                 guide = "legend",
                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})

@@

  clean_years %>%
    mutate(per = cut(per, 
                     breaks = c(0, 1, 2, Inf),
                     labels = c("1", "2", "3 or more")))  %>%
    group_by(year, per) %>%  
    # Take the median instead of mean to avoid extreme observations skewing the 
    # result.
    summarise(new_y = median(burden, na.rm = T)) %>%
    ggplot(aes(x = year,   
                      y = new_y,   
                      colour = per)) +  
    geom_line(size = 1) +
    labs(colour = "Household Size",
         x = "Year",
         y = "Median Burden",
         title = "Relationship between Burden and Household Size") +
    scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
    theme_tufte()

@@

```{r, echo = F, fig.align = 'center', fig.width = 12}
clean_years %>% 
  # Group and summarise the amount of people in each poverty group each year
  group_by(fmtincrelpovcat, year) %>%
  summarise(count = n()) %>%
  # Clean NAs
  na.omit() %>%
  # Group again by year to calculate relative percentages per year
  group_by(year) %>% 
  mutate(percent = count / sum(count),  
         # Clean the labels
         fmtincrelpovcat = factor(fmtincrelpovcat, 
                                  levels = c("1 LTE Poverty", 
                                             "2 100-150% Poverty", 
                                             "3 150-200% Poverty", 
                                             "4 200%+ Poverty"),  
                                  labels = c("Less than poverty income",   
                                             "100-150% of poverty income",   
                                             "150-200% of poverty income",    
                                             "> 200% of poverty income"),
                                  ordered = T)) %>% 
  ggplot(aes(x = year, 
             y = percent,
             fill = fmtincrelpovcat)) + 
  geom_area(alpha = 0.8) +
  # Use a nice colour palette
  scale_fill_brewer(palette = "RdYlGn", 
                    # Reverse the order of legend items to match the
                    # stacking order
                    guide = guide_legend(reverse=TRUE)) +
  labs(title = "Number of Households Relative to the Poverty Line",
       x = "Year",
       y = "Percent of Households",
       fill = "Household Income \nRelative to Poverty Income") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()
```

@@

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

@@

@@

* Housing Affordability Has Decreased

* The 2007 Housing Bubble Crisis

* Different Housing Regions Affected Differently

  * Western and Northeast Regions Vs. South and Midwest

>>>>

Combined Years:

- Dataset is split into fifteen datasets, one for each odd year between 1985 and 2013
- To standardize the data used data from 2003 - 2013
- Reorder the columns to exactly match that of our base (2003 onward)
- Combine every list of datasets into single dataset

Clean Data:

- Initally removed NAs, expressed column in read_csv
- Set column names to lowercase, removed quotations
- Created column for owning and renting data (1= own, 2= rent)
- Write to csv (same for combined years)

@@

## Variables According to the HUD Document Data Set

@@

@@

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),
                     breaks=c(0,20000,40000,60000),
                     labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

@@

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F }

  cleanyearsmonthly %>%  
  group_by_("year","region", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

@@

map_df %>% 
  ggplot() +  
  geom_polygon(aes(x=long, 
                   y = lat, 
                   fill = zinc2_median, 
                   group = group)) + 
  coord_fixed(1.3) +
  theme_tufte(base_size = 20) +
  scale_x_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  scale_y_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  labs(fill = "Median income") +
  scale_fill_distiller(limits = c(25000, 42000), 
                                 breaks = sort(unique(map_df$zinc2_median),
                                               decreasing = T),
                                 palette = "Greens", 
                                 direction = 1,
                                 guide = "legend",
                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})

@@

  clean_years %>%
    mutate(per = cut(per, 
                     breaks = c(0, 1, 2, Inf),
                     labels = c("1", "2", "3 or more")))  %>%
    group_by(year, per) %>%  
    # Take the median instead of mean to avoid extreme observations skewing the 
    # result.
    summarise(new_y = median(burden, na.rm = T)) %>%
    ggplot(aes(x = year,   
                      y = new_y,   
                      colour = per)) +  
    geom_line(size = 1) +
    labs(colour = "Household Size",
         x = "Year",
         y = "Median Burden",
         title = "Relationship between Burden and Household Size") +
    scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
    theme_tufte(base_size = 20)

@@

```{r, echo = F, fig.align = 'center', fig.width = 12}
clean_years %>% 
  # Group and summarise the amount of people in each poverty group each year
  group_by(fmtincrelpovcat, year) %>%
  summarise(count = n()) %>%
  # Clean NAs
  na.omit() %>%
  # Group again by year to calculate relative percentages per year
  group_by(year) %>% 
  mutate(percent = count / sum(count),  
         # Clean the labels
         fmtincrelpovcat = factor(fmtincrelpovcat, 
                                  levels = c("1 LTE Poverty", 
                                             "2 100-150% Poverty", 
                                             "3 150-200% Poverty", 
                                             "4 200%+ Poverty"),  
                                  labels = c("Less than poverty income",   
                                             "100-150% of poverty income",   
                                             "150-200% of poverty income",    
                                             "> 200% of poverty income"),
                                  ordered = T)) %>% 
  ggplot(aes(x = year, 
             y = percent,
             fill = fmtincrelpovcat)) + 
  geom_area(alpha = 0.8) +
  # Use a nice colour palette
  scale_fill_brewer(palette = "RdYlGn", 
                    # Reverse the order of legend items to match the
                    # stacking order
                    guide = guide_legend(reverse=TRUE)) +
  labs(title = "Number of Households Relative to the Poverty Line",
       x = "Year",
       y = "Percent of Households",
       fill = "Household Income \nRelative to Poverty Income") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)
```

## Income Compared to Area's Median by House TypeS

```{r, echo = F, fig.align = 'center', fig.width = 12, message =F}

clean_years %>% filter(fmtstructuretype != ".") %>% 
      # Use NSE
      group_by_("year", "fmtstructuretype") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(x = "year",   
                        y = "new_y",   
                        colour = "fmtstructuretype")) +  
  geom_line(size = 1) + 
  labs(colour = "Structure Type",
       y="Income percentage of area median income",
       x = "Year") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

## Burden Based on Median Housing Cost with Assistance

```{r, echo = F, fig.align = 'center', fig.width = 12, message= F, warning=F}

assisted = clean_years %>% filter(!is.na(assisted)) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
      ungroup %>% 
      mutate(assisted = as.factor(assisted))
ggplot(assisted, aes(y = new_y,   
                  x = fmtcostmedrelamicat,
                  fill = assisted, group = assisted)) +  
  geom_bar(
           stat="identity") + 
  labs(y="Burden",
       x = "",
       fill="Assistance") + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer("Assistance", 
                    labels = c("Without Assistance", 
                               "With Assistance"),
                    palette = "Set2")  +
  coord_flip() +
  theme_tufte(base_size = 20)

@@

```{r, echo = F, fig.align = 'center', fig.width = 12, message= F}

clean_years %>%
  mutate(region = as.character(region)) %>% 
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) + 
  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
       title = "Median Income Adjusted \nfor the Number of People by Region") + 
  scale_colour_brewer(palette = "Dark2", 
                      name = "Region", 
                      labels = regionNames) +
  theme_tufte(base_size = 20)
```

- Adjusted income per persons
- Bracketed by HUD indexes (0.70 for 1 person, 0.80 for 2, 1.00 for 4)
- Spike around 2010 associated with financial crisis

## Median Income Adjusted for Number of Bedrooms By Region

clean_years %>%
  mutate(region = as.character(region)) %>% 
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(ablmed, na.rm = T)) %>%
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) +
  labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)") + 
  scale_colour_brewer(palette = "Dark2",
                      name = "Region", 
                      labels = regionNames) +
  theme_tufte(base_size = 20)

@@

Adjusted income per bedroom

Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)

More discrepancy in burden levels between the Midwest/South and Northeast/West

  - Larger homes in the Midwest/South

Purhasing parity shoudl be higher in midwest/south, not expressed in data

@@

Housing Affordability Has Decreased

The 2007 Housing Bubble Crisis and it's Consequences

Census Regions Differences

  - Western and Northeast Regions Vs. South and Midwest
